---
output: 
  github_document
---

# rSharp


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  message = FALSE,
  warning = FALSE,
  out.width = "100%"
)
```


<!-- README.md is generated from README.Rmd. Please edit that file -->


The **RSharp** R package provides access to .NET libraries from R. It allows
to create .NET objects, access their fields, and call their methods.

This package is based on the [rClr](https://github.com/rdotnet/rClr) package and
utilizes some of its code base.

The package utilizes two code bases - the R code which is exposed to the user, 
and the C++/C# code that communicates with the .NET libraries. The end user of 
the package will only interact with the R code, and pre-build C++/C# libraries are
supplied with the package. See section [Installation](#installation) for instructions on how to install the package.

Advanced users and C++/C# developers might want to build the C++/C# code from source.
For this, follow the instructions in section [Build](#build).

## Installation 

### Windows

You will need the latest Microsoft Visual C++ Redistributable for Visual Studio 2015, 2017 and 2019 (you can get it from [here](https://support.microsoft.com/en-us/help/2977003/the-latest-supported-visual-c-downloads) 
and .NET 8 runtime (you can get it from [here](https://dotnet.microsoft.com/download/dotnet/8.0/runtime)).

You can install the package from GitHub with pre-build binaries by running:

```{r, eval=FALSE}
install.packages("remotes")
remotes::install_github("Open-Systems-Pharmacology/rSharp@*release")
```

Get the latest development version with:

```{r, eval=FALSE}
remotes::install_github("Open-Systems-Pharmacology/rSharp")
```

Alternatively, download the [latest release](https://github.com/Open-Systems-Pharmacology/rSharp/releases) of the package and install it locally.

### Ubuntu

Run the following commands to install the required dependencies:

```
sudo apt-get install dotnet-runtime-8.0 libcurl4-openssl-dev libssl-dev libxml2-dev 
sudo apt-get install libfontconfig1-dev libharfbuzz-dev libfribidi-dev
sudo apt-get install libfreetype6-dev libpng-dev libtiff5-dev libjpeg-dev
sudo ln -s /usr/lib/x86_64-linux-gnu/libdl.so.2 /usr/lib/x86_64-linux-gnu/libdl.so
```
Symlink libdl.so.2 to libdl.so

```
sudo ln -s /usr/lib/x86_64-linux-gnu/libdl.so.2 /usr/lib/x86_64-linux-gnu/libdl.so
```

Install the package from GitHub with pre-build binaries by running:

```{r, eval=FALSE}
install.packages("remotes")
remotes::install_github("Open-Systems-Pharmacology/rSharp@*release")
```

Get the latest development version with:

```{r, eval=FALSE}
remotes::install_github("Open-Systems-Pharmacology/rSharp")
```

## Build

### Windows

To build the C++/C# libraries from source, make sure the ```Desktop development with C++``` workload is installed in Visual Studio.

Create an environment variable ```R_INSTALL_PATH``` and set the value to the path where R is installed

```
set R_INSTALL_PATH = "C:\Program Files\R\R-4.3.3"
```

Start Visual Studio and open the ```rSharp.sln``` solution file and build the solution.
The build process will automatically copy the required files to the `rSharp/inst/lib` folder.

Then start your preferred R environment and run the following R commands from within the ```rSharp``` folder

```
devtools::install()
```



### Ubuntu

Set up .NET SDK and `nuget` (the latter to get the required dependencies)

Optionally set up to build the binaries
```
sudo apt-get install dotnet-sdk-8.0
sudo apt-get install nuget
```

Navigate to the ```rSharp\shared``` directory and run
```
make
```

Then change back to the ```rSharp``` directory start your preferred R environment and run the following R commands
```
devtools::install()
```

## User guide

Examples of interacting with .NET assemblies using this package are detailed in `vignette('user-guide')`. Some useful tips around using the package are available in
the `vignette('knowledge-base')`.

## Code of conduct

Everyone interacting in the Open Systems Pharmacology community (codebases, issue trackers, chat rooms, mailing lists etc...) is expected to follow the Open Systems Pharmacology [code of conduct](https://github.com/Open-Systems-Pharmacology/Suite/blob/master/CODE_OF_CONDUCT.md).

## Contribution

We encourage contribution to the Open Systems Pharmacology community. Before getting started please read the [contribution guidelines](https://github.com/Open-Systems-Pharmacology/Suite/blob/master/CONTRIBUTING.md). If you are contributing code, please be familiar with the [coding standards](https://github.com/Open-Systems-Pharmacology/Suite/blob/master/CODING_STANDARDS_R.md).

## License

The `{RSharp}` package is released under the [GPLv2 License](LICENSE).

All trademarks within this document belong to their legitimate owners.
